Factors determining how catchments respond to forest cover change. Re-analysing global data sets.

### 2022-07-29 paper/Forest_and_Water/Forest_and_Water.tex
first draft response review completed

There has been an long and on-going discussion in the hydrological literature around the impact of forests on streamflow (Andréassian, 2004; Brown et al., 2013, 2005; Filoso et al., 2017; Jackson et al., 2005; Zhang et al., 2017). The historic work highlights a general consensus that if forest areas increase, streamflow decreases and vice-versa. The most dramatic result in relation to this, is Figure 5 in Zhang et al. (2011) indicating (for Australian catchments) a 100\% decrease in streamflow for catchments with 100\% forest cover. However, on the other end of the spectrum, for three French catchments (Cosandey et al., 2005), there was no change in streamflow characteristics in two of the catchments after deforestation.

@@

Several review papers have summarized the plethora of forestation and deforestation studies across the globe, in relation to paired watershed studies (Bosch and Hewlett, 1982; Brown et al., 2005), related to reforestation in particular (Filoso et al., 2017), and more generally (Jackson et al., 2005; Zhang et al., 2017). These studies aim to generalize the individual findings and to identify if there are global trends or relationships that can be developed. The most recent reviews (Filoso et al., 2017; Zhang et al., 2017) developed an impressive global database of catchment studies in relation to changes in streamflow due to changes in forest cover. The Zhang et al. (2017) dataset, which covers over 312 studies, is described in terms of the change in streamflow as a result of the change in forest cover, where studies related to both forestation (increase in forest cover) and deforestation (decrease in forest cover) were included. In contrast, the paper by Filoso et al. (2017) focused primarily on reforestation, and covered an equally impressive database of 167 studies using a systematic review. In this case the collected data is mostly coded as count data and only a subset of 37 studies was analysed for actual water yield change. There is some overlap between the two data sets, but there are also some studies unique to both sets.

@@

@@

Building on the analyses by Zhang et al. (2017) and Filoso et al. (2017), and combining their conclusions, the main hypothesis to test is that the change in streamflow is impacted by the change in forest cover. However, this change is is potentially modulated by the area under consideration (affecting the length of the flowpaths Zhou et al. (2015)), the length of the study (c.f. Jackson et al. (2005); Filoso et al. (2017)) and the climate (as indicated by either E0/Pa or latitude and longitude Filoso et al. (2017); Zhou et al. (2015)).

@@

\begin{longtable}[]{@{}lll@{}}
\toprule
Factor & Abbreviation & Definition \\
\midrule
\endhead
forest type & CF & coniferous forest \\
& BF & broadleaf forest \\
& MF & mixed forest \\
hydrological regime & RD & rain dominated \\
& SD & snow dominated \\
climate type & EL & energy limited \\
& WL & water limited \\
& EQ & equitant \\
precipitation data type & OB & observed \\
& SG & spatial gridded \\
& MD & modelled \\
assessment technique & PWE & paired watershed experiment \\
& QPW & quasi-paired watershed experiment \\
& HM & hydrological modelling \\
& EA & elastictity analysis \\
& SH & combined use of statistical methods \\
& & and hydrographs \\
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:tabmodel1} Summary results of the first regression model predicting change in streamflow from change in forest cover and accounting for the direction of the change. The first three rows relate to the model using the original data base from Zhang et al.~(2017). The bottom three rows are the results of the model including the new data. Clearly there is no major change arising from the additional data.}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endhead
\textbf{(Intercept)} & 6.24 & 5.19 & 1.2 & 0.23 \\
\textbf{DeltaF\_perc\_pos} & 0.46 & 0.08 & 5.71 & 0 \\
\textbf{Forest\_SignIncrease} & -26.95 & 5.4 & -4.99 & 0 \\
\textbf{(Intercept)} & 7.79 & 5.01 & 1.56 & 0.12 \\
\textbf{DeltaF\_perc\_pos} & 0.43 & 0.08 & 5.64 & 0 \\
\textbf{Forest\_SignIncrease} & -34.65 & 4.95 & -7 & 0 \\
\bottomrule
\end{longtable}

The overall variance explained in this model (equation \eqref{eq:eq3}) is not high with an adjusted \emph{r\textsuperscript{2}} of 0.24, it generally supports the hypothesized relationship between the change in forest cover and the change in flow. The model suggests that for every 1\% change in forest cover, on the average, the flow changes 0.45\%. However the change in flow is different for forest cover decreases compared to forest cover increases. In fact, forest cover increases decrease flow by 29\% less than a similar decrease in forest cover causes flow to increase. So roughly speaking, a 1\% forest cover increase on the average decreases flow by \((1 - 0.29)*0.45\%\), while a the percentage forest cover decrease will increase flow by 0.45\%.

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-model3} Results of the model based on the complete dataset and including Latitude and Longitude}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endhead
\textbf{(Intercept)} & 9.88 & 5.42 & 1.82 & 0.07 \\
\textbf{DeltaF\_perc\_pos} & 0.44 & 0.08 & 5.68 & 0 \\
\textbf{Forest\_SignIncrease} & -35.72 & 5.21 & -6.86 & 0 \\
\textbf{Latitude} & -0.1 & 0.09 & -1.17 & 0.24 \\
\textbf{Longitude} & -0.02 & 0.03 & -0.62 & 0.54 \\
\bottomrule
\end{longtable}

There appears to be no significant gradient in either latitude or longitude (Table \ref{tab:out-model3}), suggesting that the distribution of the catchments across the globe has little influence. The total explaining power of the model is still low with an adjusted \emph{r\textsuperscript{2}} of 0.24 suggesting further factors influencing the change in streamflow that are currently not included in the model.

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-model4} Results of the model including the dryness index}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endhead
\textbf{(Intercept)} & 4.2 & 5.86 & 0.72 & 0.47 \\
\textbf{DeltaF\_perc\_pos} & 0.44 & 0.08 & 5.7 & 0 \\
\textbf{Forest\_SignIncrease} & -35.37 & 4.97 & -7.12 & 0 \\
\textbf{Dryness} & 2.76 & 2.5 & 1.11 & 0.27 \\
\bottomrule
\end{longtable}

Similar to \emph{E0} or \emph{Pa\_mm}, the results from this model (equation \eqref{eq:eq5} and Table \ref{tab:out-model4}) interestingly indicate no impact of dryness on the change in streamflow as a function of the change in forest cover change. This might seem suprising in light of earlier reported results (Filoso et al., 2017; Zhang et al., 2017). In this case, the evidence is highly doubtful (p = 0.27). However, it is very well possible that there is a further interaction in the data with other variables or unknown variables that this simpler version of the model cannot identify. This is partly evidenced by the fact that the overall variance explained is still low, with an adjusted \emph{r\textsuperscript{2}} of 0.24. As indicated in the methods, we retain Dryness in further models as an indicator of climate for the catchments.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1667}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4167}}@{}}
\caption{\label{tab:drytable} catchments for which the dryness index \textgreater{} 4}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
Number
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Latitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Longitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Catchment name
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
Number
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Latitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Longitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Catchment name
\end{minipage} \\
\midrule
\endhead
76 & 34.67 & -111.7 & Beaver Creek, AZ \#3-2 \\
90 & 36.4 & -120.4 & Cantua \\
226 & 32.74 & -111.5 & Natural Drainages, Ariz.,
U.S.A, C \\
225 & 32.74 & -111.5 & Natural Drainages, Ariz.,
U.S.A, A \\
295 & 34.43 & -112.3 & White Spar, Ariz., U.S.A, B \\
356 & -25.75 & 28.23 & Queens river \\
\bottomrule
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-modelArea} Results of the model including Area and the dryness index}\tabularnewline
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endfirsthead
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule
\endhead
\textbf{(Intercept)} & 11.15 & 6.73 & 1.66 & 0.1 \\
\textbf{DeltaF\_perc\_pos} & 0.33 & 0.09 & 3.49 & 0 \\
\textbf{Forest\_SignIncrease} & -33.52 & 5.02 & -6.67 & 0 \\
\textbf{Dryness} & 3.52 & 2.51 & 1.4 & 0.16 \\
\textbf{log10(Area\_km2)} & -3.23 & 1.56 & -2.07 & 0.04 \\
\bottomrule
\end{longtable}

The results of this model (Equation \eqref{eq:eq6}) indicate there is at least some evidence (p = 0.04) that there is a reduction in the effect of forest cover change on streamflow related to log10 (Area (km\textsuperscript{2})) (Table \ref{tab:out-modelArea}). In fact, the results suggests that for every additional 10 km\textsuperscript{2} in catchment size the mean change in flow reduces by 3.2\%. Another interesting fact to note is that with the inclusion of Area (km\textsuperscript{2}) as a variable in the model, the effect of Dryness becomes slightly more important, possibly suggesting an interaction between Dryness and Area. Including the interaction Dryness*log10(Area (km\textsuperscript{2})) in the model (Table \ref{tab:out-modelArea-int}) results in the increased evidence (p = 0) that Dryness affects the change in flow caused by changes in forest cover and that the effect of Area is only important (p = 0.77)) as an interaction with Dryness.

>>>>

There has been an long and on-going discussion in the hydrological literature around the impact of forests on streamflow (Andréassian, 2004; Brown et al., 2013, 2005; Filoso et al., 2017; Jackson et al., 2005; Zhang et al., 2017). The historic work highlights a general consensus that if forest areas increase, streamflow decreases and vice-versa. The most dramatic result in relation to this, is Figure 5 in Zhang et al. (2011) indicating (for Australian catchments) a 100\% decrease in streamflow for catchments with 100\% forest cover. However, on the other end of the spectrum, for three French catchments (Cosandey et al., 2005), there was no change in streamflow characteristics in two of the catchments after deforestation. In contrast, a more recent study in Brazil across 324 catchments, Levy et al. (2018) found a significant increase in streamflow, particular in the dry season, showing also quite a dramatic variation in responses to deforestation. Understanding the impact of

@@

Several review papers have summarized the plethora of forestation and deforestation studies across the globe, in relation to paired watershed studies (Bosch and Hewlett, 1982; Brown et al., 2005), related to reforestation in particular (Filoso et al., 2017), and more generally (Jackson et al., 2005; Zhang et al., 2017). These studies aim to generalize the individual findings and to identify if there are global trends or relationships that can be developed. The most recent reviews (Filoso et al., 2017; Zhang et al., 2017) developed an impressive global database of catchment studies in relation to changes in streamflow due to changes in forest cover. The Zhang et al. (2017) dataset, which covers over 312 studies, is described in terms of the change in streamflow as a result of the change in forest cover, where studies related to both forestation (increase in forest cover) and deforestation (decrease in forest cover) were included. In contrast, the paper by Filoso et al. (2017) focused primarily on reforestation, and covered an equally impressive database of 167 studies using a systematic review. In this case the collected data is mostly coded as count data and only a subset of 37 studies was analysed for actual water yield change. There is some overlap between the two data sets, but there are also some studies unique to both sets. The more regionally concentrated and detailed study by Levy et al. (2018) is a further independent dataset with no overlap with the other studies. However, for this study only the flow and rainfall data is available for the catchments, and the change in landcover would need to be derived again.

@@

In contrast to the basic linear regression in the earlier studies (Filoso et al., 2017; Zhang et al., 2017) and the top-down Budyko modelling (Hoek van Dijke et al., 2022; Zhou et al., 2015), the regional Brazilian Cerrado study (Levy et al., 2018) provides a carefully designed statistical approach using mixed effects modelling and Differences-in-Differences modelling focusing specifically on the effect of deforestation. The analysis specifically accounted for differences between catchments and differences due to variations in climate

@@

Building on the analyses by Zhang et al. (2017) and Filoso et al. (2017), and combining their conclusions, the main hypothesis to test is that the change in streamflow is impacted by the change in forest cover. However, this change is is potentially modulated by the area under consideration (affecting the length of the flow paths Zhou et al. (2015)), the length of the study (c.f. Jackson et al. (2005); Filoso et al. (2017)) and the climate (as indicated by either E0/Pa or latitude and longitude Filoso et al. (2017); Zhou et al. (2015)).

@@

\begin{longtable}[]{@{}lll@{}}
\toprule()
Factor & Abbreviation & Definition \\
\midrule()
\endhead
forest type & CF & coniferous forest \\
& BF & broadleaf forest \\
& MF & mixed forest \\
hydrological regime & RD & rain dominated \\
& SD & snow dominated \\
climate type & EL & energy limited \\
& WL & water limited \\
& EQ & equitant \\
precipitation data type & OB & observed \\
& SG & spatial gridded \\
& MD & modelled \\
assessment technique & PWE & paired watershed experiment \\
& QPW & quasi-paired watershed experiment \\
& HM & hydrological modelling \\
& EA & elastictity analysis \\
& SH & combined use of statistical methods \\
& & and hydrographs \\
\bottomrule()
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:tabmodel1} Summary results of the first regression model predicting change in streamflow from change in forest cover and accounting for the direction of the change. The first three rows relate to the model using the original data base from Zhang et al.~(2017). The bottom three rows are the results of the model including the new data. Clearly there is no major change arising from the additional data.}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & 13.56 & 4.99 & 2.72 & 0.01 \\
\textbf{DeltaF\_perc\_pos} & 0.36 & 0.08 & 4.59 & 0 \\
\textbf{Forest\_SignIncrease} & -43.11 & 5.46 & -7.89 & 0 \\
\textbf{(Intercept)} & 15.14 & 4.76 & 3.18 & 0 \\
\textbf{DeltaF\_perc\_pos} & 0.33 & 0.07 & 4.48 & 0 \\
\textbf{Forest\_SignIncrease} & -48.43 & 4.85 & -9.98 & 0 \\
\bottomrule()
\end{longtable}

The overall variance explained in this model (equation \eqref{eq:eq3}) is not high with an adjusted \emph{r\textsuperscript{2}} of 0.32, it generally supports the hypothesized relationship between the change in forest cover and the change in flow. The model suggests that for every 1\% change in forest cover, on the average, the flow changes 0.45\%. However the change in flow is different for forest cover decreases compared to forest cover increases. In fact, forest cover increases decrease flow by 29\% less than a similar decrease in forest cover causes flow to increase. So roughly speaking, a 1\% forest cover increase on the average decreases flow by \((1 - 0.29)*0.45\%\), while a the percentage forest cover decrease will increase flow by 0.45\%.

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-model3} Results of the model based on the complete dataset and including Latitude and Longitude}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & 16.53 & 5.04 & 3.28 & 0 \\
\textbf{DeltaF\_perc\_pos} & 0.34 & 0.07 & 4.56 & 0 \\
\textbf{Forest\_SignIncrease} & -48.41 & 4.92 & -9.84 & 0 \\
\textbf{Latitude} & -0.09 & 0.08 & -1.05 & 0.29 \\
\textbf{Longitude} & -0.03 & 0.03 & -1.07 & 0.28 \\
\bottomrule()
\end{longtable}

There appears to be no significant gradient in either latitude or longitude (Table \ref{tab:out-model3}), suggesting that the distribution of the catchments across the globe has little influence. The total explaining power of the model is still low with an adjusted \emph{r\textsuperscript{2}} of 0.33 suggesting further factors influencing the change in streamflow that are currently not included in the model.

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-model4} Results of the model including the dryness index}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & 9.49 & 5.47 & 1.73 & 0.08 \\
\textbf{DeltaF\_perc\_pos} & 0.33 & 0.07 & 4.47 & 0 \\
\textbf{Forest\_SignIncrease} & -49.86 & 4.87 & -10.23 & 0 \\
\textbf{Dryness} & 4.66 & 2.36 & 1.97 & 0.05 \\
\bottomrule()
\end{longtable}

Similar to \emph{E0} or \emph{Pa\_mm}, the results from this model (equation \eqref{eq:eq5} and Table \ref{tab:out-model4}) interestingly indicate no impact of dryness on the change in streamflow as a function of the change in forest cover change. This might seem suprising in light of earlier reported results (Filoso et al., 2017; Zhang et al., 2017). In this case, the evidence is highly doubtful (p = 0.05). However, it is very well possible that there is a further interaction in the data with other variables or unknown variables that this simpler version of the model cannot identify. This is partly evidenced by the fact that the overall variance explained is still low, with an adjusted \emph{r\textsuperscript{2}} of 0.33. As indicated in the methods, we retain Dryness in further models as an indicator of climate for the catchments.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1667}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.4167}}@{}}
\caption{\label{tab:drytable} catchments for which the dryness index \textgreater{} 4}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Number
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Latitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Longitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Catchment name
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
Number
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Latitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Longitude
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Catchment name
\end{minipage} \\
\midrule()
\endhead
76 & 34.67 & -111.7 & Beaver Creek, AZ \#3-2 \\
90 & 36.4 & -120.4 & Cantua \\
225 & 32.74 & -111.5 & Natural Drainages, Ariz.,
U.S.A, A \\
226 & 32.74 & -111.5 & Natural Drainages, Ariz.,
U.S.A, C \\
295 & 34.43 & -112.3 & White Spar, Ariz., U.S.A, B \\
356 & -25.75 & 28.23 & Queens river \\
\bottomrule()
\end{longtable}

@@

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.3611}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1806}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1389}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1528}}@{}}
\caption{\label{tab:out-modelArea} Results of the model including Area and the dryness index}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & 12.99 & 6.29 & 2.07 & 0.04 \\
\textbf{DeltaF\_perc\_pos} & 0.27 & 0.09 & 3.09 & 0 \\
\textbf{Forest\_SignIncrease} & -48.57 & 5 & -9.71 & 0 \\
\textbf{Dryness} & 5.02 & 2.38 & 2.11 & 0.04 \\
\textbf{log10(Area\_km2)} & -1.68 & 1.49 & -1.13 & 0.26 \\
\bottomrule()
\end{longtable}

The results of this model (Equation \eqref{eq:eq6}) indicate there is at least some evidence (p = 0.26) that there is a reduction in the effect of forest cover change on streamflow related to log10 (Area (km\textsuperscript{2})) (Table \ref{tab:out-modelArea}). In fact, the results suggests that for every additional 10 km\textsuperscript{2} in catchment size the mean change in flow reduces by 1.7\%. Another interesting fact to note is that with the inclusion of Area (km\textsuperscript{2}) as a variable in the model, the effect of Dryness becomes slightly more important, possibly suggesting an interaction between Dryness and Area. Including the interaction Dryness*log10(Area (km\textsuperscript{2})) in the model (Table \ref{tab:out-modelArea-int}) results in the increased evidence (p = 0) that Dryness affects the change in flow caused by changes in forest cover and that the effect of Area is only important (p = 0.51)) as an interaction with Dryness.

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.4000}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1467}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1733}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1333}}
  >{\centering\arraybackslash}p{(\columnwidth - 8\tabcolsep) * \real{0.1467}}@{}}
\caption{\label{tab:out-modelArea-int} Results of the model including an interaction between Area and the dryness index}\tabularnewline
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endfirsthead
\toprule()
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Estimate
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Std. Error
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
t value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Pr(\textgreater\textbar t\textbar)
\end{minipage} \\
\midrule()
\endhead
\textbf{(Intercept)} & 9.96 & 6.42 & 1.55 & 0.12 \\
\textbf{DeltaF\_perc\_pos} & 0.26 & 0.09 & 2.96 & 0 \\
\textbf{Forest\_SignIncrease} & -46.37 & 5.09 & -9.12 & 0 \\
\textbf{Dryness} & 7.77 & 2.71 & 2.87 & 0 \\
\textbf{log10(Area\_km2)} & 1.35 & 2.07 & 0.65 & 0.51 \\
\textbf{Dryness:log10(Area\_km2)} & -2.56 & 1.22 & -2.09 & 0.04 \\
\bottomrule()
\end{longtable}
